Liveness-Based Garbage Collection for Lazy Languages

### 2015-07-09 paper/paper.tex
figure fixed

\begin{figure}[t!]
  \begin{tabular}{cc}
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(40,12)
  \putnode{t0}{origin}{5}{0}{\var{\Df{\length}{1}}}
  \putnode{t1}{t0}{15}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \putnode{t2}{t1}{15}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \ncline{t1}{t2}
  \putnode{l0}{t1}{7}{2}{\clazy}
  \nccurve[angleA=45, angleB=135, ncurv=4, nodesep=-.5]{t1}{t1}
  \putnode{l1}{t1}{0}{10}{\acdr/\clazy}
  \nccurve[angleA=45, angleB=135, ncurv=3, nodesep=-.5]{t2}{t2}
  \putnode{l2}{t2}{0}{10}{\clazy}
\end{pspicture}
&
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(40,12)
  \putnode{t0}{origin}{5}{0}{\var{\Df{\length}{1}}}
  \putnode{t1}{t0}{15}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \nccurve[angleA=45, angleB=135, ncurv=4, nodesep=-.5]{t1}{t1}
  \putnode{l1}{t1}{0}{10}{\acdr}
\end{pspicture}
\\
(a) & (b) \\
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(40,12) %\psframe(0,-5)(40,12)
  \putnode{t0}{origin}{5}{0}{\Lanv{9}{\pa}}
  \putnode{t1}{t0}{9}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \putnode{t2}{t1}{12}{0}{\pscirclebox{\mbox{\ \ \ \ }}}
  \putnode{t3}{t2}{12}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \ncline{t1}{t2}
  \putnode{l0}{t1}{5}{2}{\clazy}
  \ncline{t2}{t3}
  \putnode{l1}{t2}{5}{2}{\bcar}
  \nccurve[angleA=55, angleB=125, ncurv=3, nodesep=-.5]{t3}{t3}
  \putnode{l2}{t3}{0}{10}{\acdr}
\end{pspicture}
&
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(40,12) %\psframe(0,-5)(40,12)
  \putnode{t0}{origin}{5}{0}{\Lanv{9}{\pa}}
  \putnode{t1}{t0}{9}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \putnode{t3}{t1}{12}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \ncline{t1}{t3}
  \putnode{l1}{t1}{5}{2}{\bcar}
  \nccurve[angleA=55, angleB=125, ncurv=3, nodesep=-.5]{t3}{t3}
  \putnode{l2}{t3}{0}{10}{\acdr}
\end{pspicture}
\\
(c) & (d)  \end{tabular}
\\ \\ \\ \\
\caption{(a) The  grammar rules for $\langle  \Df{\length}{1} \rangle$
  converted into an automaton, and simplified in (b). (c) and (d) show
  the   automata  for   \Lanv{9}{\pa}.   }\label{fig:example-automata}
\figrule
\end{figure}

@@

%% There are a couple of  issues of concern.  The garbage collection time
%% is larger in the case of LGC for some programs. The reason is that the
%% cost of consulting the liveness DFA may outweigh the combined benefits
%% of  fewer   garbage  collections   and  fewer  markings   per  garbage
%% collection.   The  other issue  is  illustrated  by  the program  {\tt
%%   lambda}    (memory   usage  shown   separately    in
%% Figure~\ref{fig:memory-usage-lambda}).  As can be seen from the table
%% in Figure~\ref{fig:experimental-results}, the  number of
%% touched cells\footnote{These  are the  cells which are  visited during
%%   marking phase, often  more than once due to  sharing.} in this
%% example is much higher for
%% LGC.  This  increase is due to  excessive sharing among  heap nodes in
%% this program.  Note  that a node re-visited because  of sharing is not
%% explored  any   further  during  a  RGC   collection.   However,  this
%% curtailment cannot happen in  LGC because of the possibility that the
%% node,  re-visited in a  different liveness  state, may  mark a  set of
%% cells different from the earlier visit.
%% }
\begin{figure*}[t]
\footnotesize

\centering

\scalebox{.9}{\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
\hline
{Program} & {\sf sudoku} & {\sf lcss} & {\sf gc\_bench} &  {\sf
  nperm} &  {\sf fibheap} & {\sf knightstour} &
{\sf treejoin} & {\sf nqueens} & {\sf lambda}\\
\hline
\hline
{Time (msec)}&106.56 &2.19 &0.32 &1.16 &2.4 &3.05 &2.61 &0.71&8.68 \\
{DFA size} &4206 &726 &258 &526 &675 &922 &737 &241 & 694\\
{Precision(\%)} &98.4&98.8&99.9&87.1&100&94.3&99.6&98.8&79.7\\
\hline
\end{tabular}}

\centerline{(a)}

\medskip

\scalebox{0.87}{\begin{tabular}{| c | r | r | r | r | r | r  |  r | r |
r | r | r | r | r |}
\hline
    & \multicolumn{2}{c|}{\# Collected} & \multicolumn{2}{c|}{\#
Touched}
  & \multicolumn{2}{c|}{} & \multicolumn{2}{c|}{} &
\multicolumn{2}{c|}{} & \multicolumn{2}{c|}{GC time} \\

                            &   \multicolumn{2}{c|}{cells per GC}
                            &   \multicolumn{2}{c|}{cells per GC}
                            &   \multicolumn{2}{c|}{\#GCs}
                            &   \multicolumn{2}{c|}{MinHeap}
                            &   \multicolumn{2}{c|}{Avg. Drag}
                            &   \multicolumn{2}{c|}{(sec)} \\
\cline{2-13}
{Program}    &
RGC & LGC & RGC & LGC  & RGC & LGC  &   RGC & LGC & RGC & LGC & RGC &
LGC \\
\hline
\hline
    {\sf   sudoku}  &381 &1107 &1268 &281 &169 &159 & 1346  &338 &527 &5
&.008 & .026 \\
    {\sf  lcss } & 46522 &51101 &6216 &1363&30&24& 52301  &1701 &5147
&588 &.007 & .008 \\
    {\sf   gc\_bench}  &129179 &131067 &1894 &4&34&4& 131071   &6 &16970
 &4 &.13 & .0002 \\
     {\sf  nperm}  & 47586  &174478 &201585 &60882&710&386& 202597  &37507
&171878 &76618 &0.250 & .154  \\
    {\sf  fibheap} &249502  &251525 &5555 &2997&1002&975& 254520  &13558
&78720 &0 &1.185 & 5.457  \\
    {\sf  knightstour}  &2593 &314564 &907502 &319299&1161&10&508225
&307092 &206729 &82112 &464.902 & 14.124  \\
    {\sf  treejoin} & 288666  &519943 &297570 &5547&116&81& 525488  &7150
&212653 &1954 &4.335 & 4.4 \\
    {\sf   nqueens} & 283822 &1423226 &2133001 &584143&46&9& 1819579
&501093 &521826 &39465 &70.314 & 24.811 \\
    {\sf   lambda}  &205 & 556 &2072&90345 &775 &750&966 & 721  &303 &95
&.179 &1.01  \\
%    {\sf   fft }& &  & & &&&& ???? &???? &????? &????? &???? & ????
\\
%    {\sf  nboyer} & &  & &   & &  & & &  \\
%    {\sf   circsim?} & &  & &   & & & & &   \\
 \hline
\end{tabular}}

\include{autotable}

\centerline{(b)}
\vskip -0.75mm
\caption{Experimental results comparing RGC and LGC. Table (a) gives
  data related to liveness analysis, and (b) gives garbage collection
  data. The average drag time is measured in allocated cons cells.}
\label{fig:experimental-results}
\normalsize
\end{figure*}

>>>>

\begin{figure}[t!]
  \begin{tabular}{cc}
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(40,12)
  \putnode{t0}{origin}{5}{0}{\var{\Df{\length}{1}}}
  \putnode{t1}{t0}{15}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \putnode{t2}{t1}{15}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \ncline{t1}{t2}
  \putnode{l0}{t1}{7}{2}{\clazy}
  \nccurve[angleA=45, angleB=135, ncurv=4, nodesep=-.5]{t1}{t1}
  \putnode{l1}{t1}{0}{10}{\acdr/\clazy}
  \nccurve[angleA=45, angleB=135, ncurv=3, nodesep=-.5]{t2}{t2}
  \putnode{l2}{t2}{0}{10}{\clazy}
\end{pspicture}
&
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(40,12)
  \putnode{t0}{origin}{5}{0}{\var{\Df{\length}{1}}}
  \putnode{t1}{t0}{15}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \nccurve[angleA=45, angleB=135, ncurv=4, nodesep=-.5]{t1}{t1}
  \putnode{l1}{t1}{0}{10}{\acdr}
\end{pspicture}
\\
(a) & (b) \\
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(40,12) %\psframe(0,-5)(40,12)
  \putnode{t0}{origin}{5}{0}{\Lanv{9}{\pa}}
  \putnode{t1}{t0}{9}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \putnode{t2}{t1}{12}{0}{\pscirclebox{\mbox{\ \ \ \ }}}
  \putnode{t3}{t2}{12}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \ncline{t1}{t2}
  \putnode{l0}{t1}{5}{2}{\clazy}
  \ncline{t2}{t3}
  \putnode{l1}{t2}{5}{2}{\bcar}
  \nccurve[angleA=55, angleB=125, ncurv=3, nodesep=-.5]{t3}{t3}
  \putnode{l2}{t3}{0}{10}{\acdr}
\end{pspicture}
&
\psset{unit=1mm,nodesep=0mm,labelsep=0.5mm}
\begin{pspicture}(0,-5)(40,12) %\psframe(0,-5)(40,12)
  \putnode{t0}{origin}{5}{0}{\Lanv{9}{\pa}}
  \putnode{t1}{t0}{9}{0}{\pscirclebox{\mbox{\ \ \ \ }}} \hspace{5mm}
  \putnode{t3}{t1}{12}{0}{\pscirclebox[doubleline=true]{\mbox{\ \ \ \ }}}
  \psset{arrows=->}
  \ncline{t0}{t1}
  \ncline{t1}{t3}
  \putnode{l1}{t1}{5}{2}{\bcar}
  \nccurve[angleA=55, angleB=125, ncurv=3, nodesep=-.5]{t3}{t3}
  \putnode{l2}{t3}{0}{10}{\acdr}
\end{pspicture}
\\
(c) & (d)  \end{tabular}
\caption{(a) The  grammar rules for $\langle  \Df{\length}{1} \rangle$
  converted into an automaton, and simplified in (b). (c) and (d) show
  the   automata  for   \Lanv{9}{\pa}.   }\label{fig:example-automata}
\figrule
\end{figure}

@@

%% There are a couple of  issues of concern.  The garbage collection time
%% is larger in the case of LGC for some programs. The reason is that the
%% cost of consulting the liveness DFA may outweigh the combined benefits
%% of  fewer   garbage  collections   and  fewer  markings   per  garbage
%% collection.   The  other issue  is  illustrated  by  the program  {\tt
%%   lambda}    (memory   usage  shown   separately    in
%% Figure~\ref{fig:memory-usage-lambda}).  As can be seen from the table
%% in Figure~\ref{fig:experimental-results}, the  number of
%% touched cells\footnote{These  are the  cells which are  visited during
%%   marking phase, often  more than once due to  sharing.} in this
%% example is much higher for
%% LGC.  This  increase is due to  excessive sharing among  heap nodes in
%% this program.  Note  that a node re-visited because  of sharing is not
%% explored  any   further  during  a  RGC   collection.   However,  this
%% curtailment cannot happen in  LGC because of the possibility that the
%% node,  re-visited in a  different liveness  state, may  mark a  set of
%% cells different from the earlier visit.
%% }
\begin{figure*}[t]
\footnotesize
\centering
%% \scalebox{.9}{\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
%% \hline
%% {Program} & {\sf sudoku} & {\sf lcss} & {\sf gc\_bench} &  {\sf
%%   nperm} &  {\sf fibheap} & {\sf knightstour} &
%% {\sf treejoin} & {\sf nqueens} & {\sf lambda}\\
%% \hline
%% \hline
%% {Time (msec)}&106.56 &2.19 &0.32 &1.16 &2.4 &3.05 &2.61 &0.71&8.68 \\
%% {DFA size} &4206 &726 &258 &526 &675 &922 &737 &241 & 694\\
%% {Precision(\%)} &98.4&98.8&99.9&87.1&100&94.3&99.6&98.8&79.7\\
%% \hline
%% \end{tabular}}
%% \centerline{(a)}
%% \medskip
%% \scalebox{0.87}{\begin{tabular}{| c | r | r | r | r | r | r  |  r | r |
%% r | r | r | r | r |}
%% \hline
%%     & \multicolumn{2}{c|}{\# Collected} & \multicolumn{2}{c|}{\#
%% Touched}
%%   & \multicolumn{2}{c|}{} & \multicolumn{2}{c|}{} &
%% \multicolumn{2}{c|}{} & \multicolumn{2}{c|}{GC time} \\
%%                             &   \multicolumn{2}{c|}{cells per GC}
%%                             &   \multicolumn{2}{c|}{cells per GC}
%%                             &   \multicolumn{2}{c|}{\#GCs}
%%                             &   \multicolumn{2}{c|}{MinHeap}
%%                             &   \multicolumn{2}{c|}{Avg. Drag}
%%                             &   \multicolumn{2}{c|}{(sec)} \\
%% \cline{2-13}
%% {Program}    &
%% RGC & LGC & RGC & LGC  & RGC & LGC  &   RGC & LGC & RGC & LGC & RGC &
%% LGC \\
%% \hline
%% \hline
%%     {\sf   sudoku}  &381 &1107 &1268 &281 &169 &159 & 1346  &338 &527 &5
%% &.008 & .026 \\
%%     {\sf  lcss } & 46522 &51101 &6216 &1363&30&24& 52301  &1701 &5147
%% &588 &.007 & .008 \\
%%     {\sf   gc\_bench}  &129179 &131067 &1894 &4&34&4& 131071   &6 &16970
%%  &4 &.13 & .0002 \\
%%      {\sf  nperm}  & 47586  &174478 &201585 &60882&710&386& 202597  &37507
%% &171878 &76618 &0.250 & .154  \\
%%     {\sf  fibheap} &249502  &251525 &5555 &2997&1002&975& 254520  &13558
%% &78720 &0 &1.185 & 5.457  \\
%%     {\sf  knightstour}  &2593 &314564 &907502 &319299&1161&10&508225
%% &307092 &206729 &82112 &464.902 & 14.124  \\
%%     {\sf  treejoin} & 288666  &519943 &297570 &5547&116&81& 525488  &7150
%% &212653 &1954 &4.335 & 4.4 \\
%%     {\sf   nqueens} & 283822 &1423226 &2133001 &584143&46&9& 1819579
%% &501093 &521826 &39465 &70.314 & 24.811 \\
%%     {\sf   lambda}  &205 & 556 &2072&90345 &775 &750&966 & 721  &303 &95
%% &.179 &1.01  \\
%% %    {\sf   fft }& &  & & &&&& ???? &???? &????? &????? &???? & ????
%% \\
%% %    {\sf  nboyer} & &  & &   & &  & & &  \\
%% %    {\sf   circsim?} & &  & &   & & & & &   \\
%%  \hline
%% \end{tabular}}
%% \centerline{(b)}

\include{autotable}
\vskip -0.75mm
\caption{Experimental results comparing RGC and LGC. Table (a) gives
  data related to liveness analysis, and (b) gives garbage collection
  data. The average drag time is measured in allocated cons cells.}
\label{fig:experimental-results}
\normalsize
\end{figure*}
